---
title: "Titanic_Vedant_Feat"
author: "Vedant"
date: "June 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd('C:/vedant/StatisticalLearning/Titanic_DataSet')
train <- read.csv("train.csv")
test <- read.csv("test.csv")

# Examine structure of dataframe
str(train)

# Look at number of people who survived
table(train$Survived)
prop.table(table(train$Survived))
# Create new column in test set with our prediction that everyone dies
test$Survived <- rep(0, 418)
# Create submission dataframe and output to file
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "theyallperish.csv", row.names = FALSE)

```
First prediction is that most people died

```{r}
# Look at gender patterns
summary(train$Sex)
prop.table(table(train$Sex, train$Survived))
## rowise proprtion can be determined as below

prop.table(table(train$Sex, train$Survived),1)


# Create new column in test set with our prediction that everyone dies
test$Survived <- 0
# Update the prediction to say that all females will survive
test$Survived[test$Sex == 'female'] <- 1
# Create submission dataframe and output to file
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "gendermodel.csv", row.names = FALSE)

# Look at age patterns
summary(train$Age)

train$Child[train$Age < 18] <- 1

aggregate(Survived ~ Child + Sex, data=train, FUN=sum)
aggregate(Survived ~ Child + Sex, data=train, FUN=length)
aggregate(Survived ~ Child + Sex, data=train, FUN=function(x) {sum(x)/length(x)})

```


```{r}

# Look at class and fare patterns
train$Fare2 <- '30+'
train$Fare2[train$Fare < 30 & train$Fare >= 20] <- '20-30'
train$Fare2[train$Fare < 20 & train$Fare >= 10] <- '10-20'
train$Fare2[train$Fare < 10] <- '<10'

aggregate(Survived ~ Fare2 + Pclass + Sex, data=train, FUN=function(x) {sum(x)/length(x)})


# Create new column in test set with our prediction that everyone dies
test$Survived <- 0

# Update the prediction to say that all females will survive
test$Survived[test$Sex == 'female'] <- 1
# Update once more to say that females who pay more for a third class fare also perish
test$Survived[test$Sex == 'female' & test$Pclass == 3 & test$Fare >= 20] <- 0

# Create submission dataframe and output to file
submit <- data.frame(PassengerId = test$PassengerId, Survived = test$Survived)
write.csv(submit, file = "genderclassmodel.csv", row.names = FALSE)


```



```{r}
# Install and load required packages for fancy decision tree plotting
library(rpart)
library(rattle)
library(rpart.plot)
library(RColorBrewer)

# Recreate the gender model

fit <- rpart(Survived ~ Sex, data=train, method="class")

fancyRpartPlot(fit)
# Build a deeper tree
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data=train, method="class")
# Plot it with base-R
plot(fit)
text(fit,cex=0.9)
# And then make it look better with fancyRpartPlot!
fancyRpartPlot(fit)
# Now let's make a prediction and write a submission file
Prediction <- predict(fit, test, type = "class")
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
write.csv(submit, file = "myfirstdtree.csv", row.names = FALSE)

# Let's unleash the decision tree and let it grow to the max
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data=train,
             method="class", control=rpart.control(minsplit=2, cp=0))
fancyRpartPlot(fit)

# Now let's make a prediction and write a submission file
Prediction <- predict(fit, test, type = "class")
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
write.csv(submit, file = "myfullgrowntree.csv", row.names = FALSE)

# Manually trim a decision tree
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data=train,
             method="class", control=rpart.control(minsplit=2, cp=0.005))
# new.fit <- prp(fit,snip=TRUE)$obj
# fancyRpartPlot(new.fit)

```


## Feature Engineering


```{r}
test$Survived = NA
head(test)
test = test[,c(1,12,2:11)]
train = train[,-c(13,14)]
head(test)
head(train)
whole = rbind(train,test)


```
 What is in the name
 
 
```{r}


##How can a person's title can give more insight ?
##There seems to be a fair few possibilities of patterns in this that may dig deeper than the combinations of age, gender, etc that we looked at before

whole$Name[1:2]

## Title starts with comma and ends with a fullstop
whole$Name = as.character(whole$Name)
whole$Title <- sapply(whole$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][2]})

whole$Title <- sub(' ', '', whole$Title)
table(whole$Title)
```
 Similar titles :
 
 Mme,Mrs --197+ 1
 
 mlle + miss +MS -- 2+262 +2
 
 Lady+ Dona + the countess +johnkeer = 1+1+1+1
 
 Capt + Col +Don+ Dr + Major + Rev+ Sir = 1+4+1+8+2+8+1
 
 
 
 
```{r}
whole$Title[whole$Title %in% c("Mme","Mrs")]='Mrs'
whole$Title[whole$Title %in% c("Mlle","Miss","Ms")]='Miss'
whole$Title[whole$Title %in% c("Lady","Dona","the Countess","Jonkheer")]='Lady'
whole$Title[whole$Title %in% c("Capt","Col","Don","Dr","Major","Rev","Sir")]='Sir'

table(whole$Title)
```

Create a new variable called family size

```{r}
whole$FamSize = whole$SibSp+whole$Parch+1

```

Create a new variable FamilyId



```{r}
whole$Surname = sapply(whole$Name, FUN=function(x) {strsplit(x, split='[,.]')[[1]][1]})

## Same family memebers same title+ same famsize + same embark station

whole$FamilyId = paste(as.character(whole$FamSize), whole$Surname,as.character(whole$Embarked), sep="")

whole$FamilyId[whole$FamSize <= 2] <- 'Small'
table(whole$FamilyId)
famIds=(table(whole$FamilyId))
famIds=as.data.frame(famIds)

```
further filter out any small families
```{r}
## Replace fsmIds with only small families

famIds= famIds[famIds$Freq<=2,]

## Now replace all the family sizes 

whole$FamilyId[whole$FamilyId %in% famIds$Var1]='Small'
whole$FamilyId = as.factor(whole$FamilyId)
```

divide the trainig and test set once again

```{r}
train = whole[1:891,]
test = whole[892:1309,]
```

fitting using trees

```{r}
fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamSize + FamilyId,
               data=train, 
               method="class")
fancyRpartPlot(fit)
tree.pred = predict(fit, test, type = "class")
submit <- data.frame(PassengerId = test$PassengerId, Survived = tree.pred)
write.csv(submit, file = "engineeredfeaturestree.csv", row.names = FALSE)
```

## use random forest


```{r}
sapply(whole, function(x) sum(is.na(x)))
## fill the age using Anova function

Agefit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title + FamSize, data=whole[!is.na(whole$Age),], method="anova")
whole$Age[is.na(whole$Age)] <- predict(Agefit, whole[is.na(whole$Age),])

## fill the missing fare value

whole$Fare[is.na(whole$Fare)] = median(whole$Fare,na.rm = TRUE)

```


```{r}
# New factor for Random Forests, only allowed <32 levels, so reduce number
whole$FamilyID2 <- whole$FamilyId
# Convert back to string
whole$FamilyID2 <- as.character(whole$FamilyID2)
whole$FamilyID2[whole$FamSize <= 3] <- 'Small'
table(whole$FamilyID2)
# And convert back to factor
whole$FamilyID2 <- factor(whole$FamilyID2)
whole$Title = as.factor(whole$Title)
# Split back into test and train sets
train <- whole[1:891,]
test <- whole[892:1309,]
library(randomForest)
set.seed(415)
fit <- randomForest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamSize + FamilyID2,
                    data=train, importance=TRUE, ntree=2000)
# Look at variable importance
varImpPlot(fit)

pred.randomforest = predict(fit,test)
pred.randomforest[1:10]

rf_df = data.frame(PassengerId=test$PassengerId,Survived=  pred.randomforest)

write.csv(rf_df,file = "randomforestpred.csv",row.names = FALSE)

```



